---
title: Examining NY Times's COVID-19 Data
output: html_notebook
---

The NY Times publishes US-county-level counts of COVID-19 cases and deaths daily on [github](https://github.com/nytimes/covid-19-data).  This notebook combines this data with
the estimated population data from the US census to identify areas with high ratios of COVID-19 cases (aka "hotspots")

First, let's read in the COVID-19 data from the NY Times.  Each county is identified by a [FIPS code](https://en.wikipedia.org/wiki/FIPS_county_code).
```{r}
library(tidyverse)

col_types <- cols(
  date = col_date(format = ""),
  county = col_character(),
  state = col_character(),
  fips = col_integer(),
  cases = col_double(),
  deaths = col_double()
)
df_cases <- read_csv('./us-counties.csv', col_types=col_types)
```
Let's read in the population data from the US census.  It requires some munging to create a FIPS column, which is required for this data to be combined with the NY Times data,
```{r}
col_types <- cols(
  .default = col_double(),
  SUMLEV = col_character(),
  STATE = col_integer(),
  COUNTY = col_integer(),
  STNAME = col_character(),
  CTYNAME = col_character()
)

df_pop <- read_csv('./co-est2019-alldata.csv', col_types=col_types) %>%
    select( "COUNTY", "STATE", "POPESTIMATE2019") %>%
    mutate(fips=STATE*1000 + COUNTY) %>%
    select("fips", POPESTIMATE2019)
```
Merge it into the NY Times data by the shared column, fips,
```{r}
df_cases <- left_join(df_cases, df_pop, by="fips")
```
Let's add a new column that is the ratio of COVID-19 cases to each county's population.  Counties with a high ratio could be classified as a hotspot.  Note that the NYC counties (King's, Queen's, etc.) are a hotspot right now but are missing.  That's
because the NY Times has chosen to break NYC's data out separately.

Let's print out the top 100 counties with the highest ratio,
```{r}
df_cases <- df_cases %>%
    mutate(ratio = cases/POPESTIMATE2019)

df_cases %>%
    arrange(desc(date), desc(ratio)) %>%
    top_n(100)
```

Let's plot the cases vs. time a few of the hotspots.  Are any of them showing a deceleration in the number of cases?

```{r}
library(ggplot2)

plotCases <- function(f) {
    data <- df_cases %>% filter(fips==f)
    elem <- (data %>% slice(1))
    ggplot(data, aes(date, cases)) + geom_point() + ggtitle(sprintf('COVID-19 Cases vs. Date for %s, %s (FIPS=%d)',
        elem$county, elem$state, elem$fips))
}

plotCases(36119)
plotCases(36087)
plotCases(16013)
plotCases(22071)

# a few place of personal interest
plotCases(6085)
plotCases(24027)
plotCases(34041)
plotCases(41067)
plotCases(45043)
```