---
title: Examining NY Times's COVID-19 Data
output: html_notebook

---

The NY Times publishes US-county-level counts of COVID-19 cases and deaths daily on [github](https://github.com/nytimes/covid-19-data).  This notebook combines estimated population data from the US census to identify areas with high ratios of COVID-19 cases.

Read in the COVID-19 data from the nytimes:
```{r}
df_cases <- read.csv('./us-counties.csv')
```
Read in the data from the US census and merge it into NY Times data,
```{r}
df_pop <- read.csv('./co-est2019-alldata.csv')
df_pop <- df_pop[c(4:5,19)]

# compute fips.
df_pop$fips <- df_pop$STATE*1000 + df_pop$COUNTY
df_pop$STATE <- NULL
df_pop$COUNTY <- NULL

total <- merge(df_cases, df_pop, by="fips")
total$ratio <- total$cases/total$POPESTIMATE2019

```
Let's examine a few cases of interest:

```{r}
subset(total, fips==34041 & date=='2020-03-29')
subset(total, fips==24027 & date=='2020-03-29')
subset(total, fips==6085 & date=='2020-03-29')
```

# Hotspots

Let's find the counties with the highest ratio of cases to population.
```{r}

# get_hotspots( ratio, date )
get_hotspots <- function(r, d) {
    hotspots <- subset(total, ratio > r & date==d)
    hotspots$date <- NULL
    hotspots <- hotspots[order(hotspots$ratio, decreasing=TRUE),]
    hotspots$deaths <- NULL  # delete to make the output fit on the page
    return(hotspots)
}

hotspots <- get_hotspots( 0.0009, '2020-03-29')
hotspots
```

```{r}
library(ggplot2)

plotCases <- function(f, name) {
    data <- subset(df_cases, fips==f)
    ggplot(data, aes(date, cases)) + geom_point() + ggtitle(paste( 'COVID-19 Cases vs. Date for ', name))
}

#plotCases(6085)

plotCases(36119, 'Rockland, NY')
plotCases(6085, 'Santa Clara, CA')
plotCases(24027, 'Howard, MD')
plotCases(34041, 'Warren, NJ')

plotCases(22071, 'New Orleans, LA')

```